### 2025-01-17 ESC195 Lecture Notes/main.tex
jan17

Evaluating the integral:
\begin{align*}
    \int_0^\infty e^{-x} \, dx &= \lim_{b \to \infty} \int_0^b e^{-x} \, dx \\
    &= \lim_{b \to \infty} -e^{-x} \Big|_0^b \\
    &= \lim_{b \to \infty} -e^{-b} + 1 \\
    &= 1
\end{align*}
\end{example}
\section{Hyperbolic Trigonometric Functions}
\begin{definition}[Hyperbolic Sine]
    The hyperbolic sine function is defined as:
    \begin{equation} \sinh(x) = \frac{e^x - e^{-x}}{2}. \end{equation}
\end{definition}
\begin{definition}[Hyperbolic Cosine]
    The hyperbolic cosine function is defined as:
    \begin{equation}  \cosh(x) = \frac{e^x + e^{-x}}{2}. \end{equation}
\end{definition}
\paragraph{Properties} These combinations of exponential functions have properties similar to the trigonometric functions.
\paragraph{Derivatives} The derivatives of the hyperbolic functions are:
\begin{align}
    \frac{d}{dx} \sinh(x) &= \cosh(x), \\
    \frac{d}{dx} \cosh(x) &= \sinh(x).
\end{align}
\paragraph{Identities} The hyperbolic functions satisfy the following identities:
\begin{align}
    \cosh^2(x) - \sinh^2(x) &= 1, \\
    \cosh(2x) &= \cosh^2(x) + \sinh^2(x), \\
    \sinh(2x) &= 2\sinh(x)\cosh(x).
\end{align}
\paragraph{Hyperbola} The hyperbolic functions are related to the hyperbola $x^2 - y^2 = 1$. Simiar to the circle, the hyperbola can be parametrized by the hyperbolic functions (e.g. $x = \cosh(t)$, $y = \sinh(t)$).
\paragraph{Area} The area of a sector of the hyperbola is given by:
\begin{equation}
    A = t/2,
\end{equation} 
where $t$ is the angle of the sector along the parametrization $\{(x,y)\mid x = \cosh(t), y = \sinh(t)\}$.
\begin{figure}[h!]
    \begin{center}
        \begin{tikzpicture}[scale=0.8]
            % Draw the axes
            \draw[->] (-5, 0) -- (5, 0) node[right] {$x$};
            \draw[->] (0, -5) -- (0, 5) node[above] {$y$};

@@

@@

@@

    F &= \int_0^b \rho g x f(x) \cdot dx

@@

>>>>

Evaluating the integral:
\begin{align*}
    \int_0^\infty e^{-x} \, dx &= \lim_{b \to \infty} \int_0^b e^{-x} \, dx \\
    &= \lim_{b \to \infty} -e^{-x} \Big|_0^b \\
    &= \lim_{b \to \infty} -e^{-b} + 1 \\
    &= 1
\end{align*}
\end{example}
\section{Hyperbolic Trigonometric Functions}
\begin{definition}[Hyperbolic Sine]
    The hyperbolic sine function is defined as:
    \begin{equation} \sinh(x) = \frac{e^x - e^{-x}}{2}. \end{equation}
\end{definition}
\begin{definition}[Hyperbolic Cosine]
    The hyperbolic cosine function is defined as:
    \begin{equation}  \cosh(x) = \frac{e^x + e^{-x}}{2}. \end{equation}
\end{definition}
\paragraph{Properties} These combinations of exponential functions have properties similar to the trigonometric functions.
\paragraph{Derivatives} The derivatives of the hyperbolic functions are:
\begin{align}
    \frac{d}{dx} \sinh(x) &= \cosh(x), \\
    \frac{d}{dx} \cosh(x) &= \sinh(x).
\end{align}
\paragraph{Identities} The hyperbolic functions satisfy the following identities:
\begin{align}
    \cosh^2(x) + \sinh^2(x) &= 1, \\
    \cosh(2x) &= \cosh^2(x) + \sinh^2(x), \\
    \sinh(2x) &= 2\sinh(x)\cosh(x).
\end{align}
\paragraph{Hyperbola} The hyperbolic functions are related to the hyperbola $x^2 - y^2 = 1$. Simiar to the circle, the hyperbola can be parametrized by the hyperbolic functions (e.g. $x = \cosh(t)$, $y = \sinh(t)$).
\paragraph{Area} The area of a sector of the hyperbola is given by:
\begin{equation}
    A = t/2,
\end{equation} 
where $t$ is the angle of the sector along the parametrization $\{(x,y)\mid x = \cosh(t), y = \sinh(t)\}$.
\begin{figure}[h!]
    \begin{center}
        \begin{tikzpicture}[scale=0.8]
            % Draw the axes
            \draw[->] (-5, 0) -- (5, 0) node[right] {$x$};
            \draw[->] (0, -5) -- (0, 5) node[above] {$y$};

@@

\subsubsection{Hydrostatic Pressure and Force}

@@

    F = \int_0^b \rho g x f(x) \cdot dx
, which is derived by considering $dx$ as the depth a horizontal slice of the tank.
\subsubsection{Center of Mass and Moments of Inertia}
\paragraph{Properties of the Center of Mass} The center of mass of a region $R$ has the following properties:
\begin{enumerate}
    \item \textbf{Symetry}: For all axis of symmetry, the center of mass lies on the axis.
    \item \textbf{Additivity}: The center of mass of a region is the weighted average of the centers of mass of its parts. The weights are the areas of the parts.
        \bar{x} = \sum_i \frac{A_i\bar{x_i}}{A}
\end{enumerate}
\paragraph{Formula} The centroid of a uniformly-dense region $R$ bounded by $y=f(x),\,x\in[a,b],\,y\in[f(a),f(b)]$ is derived as follows:
    \intertext{We first have $x_i^\star$ and $A_i$ as follows:}
    x_i^\star = \frac{1}{2}(x_i + x_{i+1}) &\quad A_i = f(x_i) \Delta x \nonumber \\
    \intertext{For $\bar{x}$, we have:}
    A\bar{x} &\approx \sum_i x_i^\star A_i \nonumber \\
    A\bar{x} &= \lim_{n \to \infty} \sum_i x_i^\star A_i  \nonumber \\
    \bar{x} &= \frac{1}{A}\int_a^b x f(x) \, dx
    \intertext{Similarly, for $\bar{y}$:}
    A\bar{y} &= \lim_{n \to \infty} \sum_i f(x_i^\star) A_i \nonumber \\
    \bar{y} &= \frac{1}{A}\int_a^b \frac{1}{2}(f(x))^2 \, dx
\begin{theorem}[Pappus's Theorem]
    The volume of a solid of revolution is given by the product of the area of the region and the distance, $R$, traveled by the centroid of the region from the axis of rotation.
        V = 2\pi R A
\end{theorem}
\begin{example}[Eliptical Torus]
    Given an eliptical torus (tall donut) with major axis $a$ and minor axis $b$, we would like to compute the volume of the torus:

@@

        V &= 2\pi R A \\
        &= 2\pi R \cdot \pi ab = 2\pi^2 abR

@@

    This could be, alternatively, shown via the washer method and the shell method.
\subsection{Applications to Economics and Biology}
``Biologists are so bad at math'' - Prof. Davis as he proceeds to erase the board and skip to the next section.
\section{Parametric Equations and Polar Coordinates}
\subsection{Curves Defined by Parametric Equations}
\paragraph{Parametric Equations} For $t \in \mathbb{R}$, we have:
    x = x(t), \quad y = y(t)
\begin{example}[Newton's Laws of Motion]
    Given $\ddot{x} = 0$, $\ddot{y} = -g$, we have:
        x(t) &= A_1 t + A_2, \\
        y(t) &= -\frac{1}{2}gt^2 + B_1 t + B_2
    At initial conditions $t = 0$, we have $x(0) = 0$, $y(0) = 0$, $\dot x(0) = v_0 \cos(\theta)$, $\dot y(0) = v_0 \sin(\theta) - \frac{1}{2}gt^2$.
\begin{example}[Ellipse]
    Given $x = a\cos(t)$, $y = b\sin(t)$, we have:
        x^2 &= a^2\cos^2(t), \\
        y^2 &= b^2\sin^2(t)
    , which satisfies $x^2/a^2 + y^2/b^2 = 1$.
\paragraph{Intersections of Parametric Curves} Given two parametric curves $C_1,\,C_2$, defined by:
    C_1:\, x_1 &= x_1(t), \quad y_1 = y_1(t) \\
    C_2:\, x_2 &= x_2(t), \quad y_2 = y_2(t)
, we can find the intersection points by solving the system of equations:
    x_1(t) &= x_2(t) \\
    y_1(t) &= y_2(t)
    Given $x_1 = 2t+t$, $y_1 = 5-4t$, $x_2 = 3 - 5\cos(\pi t)$, $y_2 = 1 + 5\sin(\pi t)$, we have:
        2t + 1 &= 3 - 5\cos(\pi t) \\
        5 - 4t &= 1 + 5\sin(\pi t)

### 2025-01-17 MAT185 Lecture Notes/main.tex
jan17

@@

@@

\subsection{Linear combinations}
\begin{definition}[Linear Combination]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}$, then:
        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n
    is a linear combination of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$

@@

\begin{definition}[Linear independence]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}$. If the equation:
        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n = \textbf{0}
    has only the trivial solution $\alpha_1 = \alpha_2 = \ldots = \alpha_n = 0$.

    Then, $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are said to be linearly independent; otherwise, they are linearly dependent.
\end{definition}
\begin{definition}[Span]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$, then:
    \begin{align*}
        \text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n) = \left\{ \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n \, | \, \alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R} \right\}
    \end{align*}
    is the set of all linear combinations of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$
\end{definition}
\end{document}

>>>>

    \begin{align}

@@

    \end{align}

@@

\begin{definition}[Intersection of Sets]
    Let $A$ and $B$ be sets, then:
    \begin{align}
        A \cap B = \left\{ x \, | \, x \in A \, \text{and} \, x \in B \right\}
    \end{align}

@@

\begin{definition}[Union of Sets]
    Let $A$ and $B$ be sets, then:
    \begin{align}
        A \cup B = \left\{ x \, | \, x \in A \, \text{or} \, x \in B \right\}
    \end{align}

\subsection{Linear combinations}
\begin{definition}[Linear Combination]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}$, then:
    \begin{align*}
        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n
    \end{align*}
    is a linear combination of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$
\end{definition}
\begin{definition}[Linear independence]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}$. If the equation:
    \begin{align*}
        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n = \textbf{0}
    \end{align*}
    has only the trivial solution $\alpha_1 = \alpha_2 = \ldots = \alpha_n = 0$.

    Then, $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are said to be linearly independent; otherwise, they are linearly dependent.
\end{definition}
\begin{definition}[Span]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$, then:
    \begin{align}
        \text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n) = \left\{ \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n \, | \, \alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R} \right\}
    \end{align}
    is the set of all linear combinations of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$
\end{definition}
\end{document}

### 2025-01-17 MSE160 Lecture Notes/main.tex
jan17

    \paragraph{Mass} The mass of the beam is given by:
    $$ m = \rho V = \rho A L $$
    \paragraph{Deflection} The deflection of the beam is given by:
    \begin{align*} 
        \delta &= \frac{FL^3}{48EI} 
        \intertext{For a beam with a rectangular cross-section, we have:}
        \delta &= \frac{FL^3}{48E} \cdot \frac{12}{bh^3} = \frac{FL^3}{4Ebh^3}  
        \intertext{We can set $b$ proportional to $h$:}
        \delta &= \frac{FL^3}{cE} \cdot \frac{1}{A^2},\quad \text{for some constant $c$} 
        \intertext{We can then isolate for $A$, the free variable, and minimize the mass via the objective equation $m = \rho A L$:}
        A &= \sqrt{\frac{FL^3}{cE\delta}} \\
        m &= \rho L \sqrt{\frac{FL^3}{cE\delta}} = \rho L \sqrt{\frac{FL^3}{cE\delta}}
        \intertext{Arrange into the form (functional)(geometric)(material):}
        m &= \left(\frac{F}{c\delta}\right)^\frac{1}{2} \cdot \left(L^\frac{5}{2}\right)\left(\frac{\rho}{E^\frac{1}{2}}\right)
    \end{align*}
\end{example}
\paragraph{Material Peformance Index} The material performance index is given by:
    \begin{equation}
        \text{Material Performance Index (MSI)} = \frac{E^\frac{1}{2}}{\rho}
\end{equation}
\paragraph{MPI Graph} We plot $\log E$ against $\log \rho$ to get the MPI graph. 
\appendix
\includepdf[pages=-, angle=90]{EquationSheet.pdf}
\end{document}

>>>>

    \paragraph{Mass} The mass of the beam is given by:
    $$ m = \rho V = \rho A L $$
    \paragraph{Deflection} The deflection of the beam is given by:
    \begin{align*} 
        \delta &= \frac{FL^3}{48EI} 
        \intertext{For a beam with a rectangular cross-section, we have:}
        \delta &= \frac{FL^3}{48E} \cdot \frac{12}{bh^3} = \frac{FL^3}{4Ebh^3}  
        \intertext{We can set $b$ proportional to $h$:}
        \delta &= \frac{FL^3}{cE} \cdot \frac{1}{A^2},\quad \text{for some constant $c$} 
        \intertext{We can then isolate for $A$, the free variable, and minimize the mass via the objective equation $m = \rho A L$:}
        A &= \sqrt{\frac{FL^3}{cE\delta}} \\
        m &= \rho L \sqrt{\frac{FL^3}{cE\delta}} = \rho L \sqrt{\frac{FL^3}{cE\delta}}
        \intertext{Arrange into the form (functional)(geometric)(material):}
        m &= \left(\frac{F}{c\delta}\right)^\frac{1}{2} \cdot \left(L^\frac{5}{2}\right)\cdot\left(\frac{\rho}{E^\frac{1}{2}}\right)
    \end{align*}
\end{example}
\paragraph{Material Peformance Index} The material performance index is given by:
    \begin{equation}
        \text{Material Performance Index (MSI)} = \frac{E^\frac{1}{2}}{\rho}
\end{equation}
\paragraph{MPI Graph} We plot $\log E$ against $\log \rho$ to get the MPI graph. 
\paragraph{Tempered Glass} Temperglass are made to resist tension. It is done by applying a compressive stress to the surface of the glass. This is done by cooling the surface of the glass faster than the core or chemically treating the surface.
\appendix
\includepdf[pages=-, angle=90]{EquationSheet.pdf}
\end{document}
